Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2013-11-21 paper/paper.tex
mostly rewrite pill section, and start synchronizing box plot with image plot Tested everything on collins.

Our diffusion and reaction rates are shown below.  We are interested
primarily in the effect of cellular shape on the protein
oscillations, so we follow Huang's work\cite{huang2003dynamic} and do
not deviate from the wild type values used in that work.  Huang's
simulations use total MinD and MinE concentrations of $1,000/\mu m$
and $350/\mu m$, respectively, in a cylindrical cell of radius $0.5\mu
m$, and in our (non-cylindrical) cells we use the same number of
proteins per unit volume.  These concentration values are 1273 $\mu
m^{-3}$ and 445 $\mu m^{-3}$, respectively.

\begin{gather*}
  \mathcal{D}_D = \mathcal{D}_{E} = 2.5\micron^2/\text{sec}\\
  k_D^{\textrm{ADP $\rightarrow$ ATP}} = 1/\textrm{sec,  }
  k_D = 0.025 \micron /\textrm{sec}\\
  k_{dD} = 0.0015 \micron^3/ \textrm{sec,  }
  k_{de} = 0.7/\textrm{sec}\\
  k_E = 0.093 \micron^3 /\textrm{sec}.
\end{gather*}

A 3d grid is then constructed in cartesian coordinates, with a grid
spacing of .05\micron. We define our cell shapes and solve the
reaction-diffusion equations numerically to observe the time evolution
of the MinD and MinE concentrations inside the cell. \renesays{This
  last sentence sweeps a lot under the rug. How much more detail
  should be given to the computational method?}

@@

We begin with the naturally occuring pill shape.  We peice this shape
together as two hemispherical endcaps attached on either end of a
cylinder.  This shape follows the early simulations of Huang \emph{et
  al.} but differs in that we have added the end caps for a more
natural shape, expecting similar results.  We test cells of radius
$.5\micron$ and of lengths $2\micron$, $3\micron$, and $5\micron$
between both tips of the end caps. In all these models we observe
quick establishement (within a few periods) of extremely regular
oscillatory behavoir that lasts indefinitely, with periods that
closely agree with the results of Huang \emph{et al.}
Table~\ref{tab:pill-periods} shows the periods of oscillations
at different cell lengths.

\begin{table}
  \begin{tabular}{|r|c|c|c|l|}
    \hline
    Length($\mu$) & 2.50 & 3.00 & 4.00 & 5.00\\
    \hline
    Period(sec) & sim & 33 & 38 & 48 \\ \cline{2-2}
    \hline
  \end{tabular}
  \caption{Period According to Pill Length}\label{tab:pill-period}
\end{table}

This simple model is a good starting point for observing in detail the
dynamic interaction between the different stages of protein that lead
to their qualitatively distinct behavoir. Figure \ref{image-p} shows
the process in a series of frames taken from simulation animations.
Each frame is 2.5 seconds ahead of the last.

Consider Figure \ref{image-p}.  In the beginning, the cytoplasmic
nATP is roughly evenly distributed throughout the cell.  There is nADP
concentrated in the left pole and it will transform into nATP.  At 2.5
seconds we see membrane bound MinD:ATP form on the right side of the
cell.  It avoids the left side because there is a concentration of
cytoplasmic MinE on the left side and compared to the right side,
which is empty of MinE, MinD:ATP attaching to the wall on the left has
no time to collect and cluster before reacting.  One can see how as
the MinE spreads out rightward, the the clustered membrane bound
MinD:ATP stays just at the edge of it, always to the right.  Where the
two meet there is at every time step a spike in membrane bound
MinD-MinE:ATP, resulting from the MinE reaction with the MinD on the
wall.  One can see from roughly 5 seconds to 17.5 seconds the
cytoplasmic MinE chases in a way the MinD towards the right, reacting
with it as it clusters on the wall.  Upon reaching the right side, the
MinD is in a way trapped, and the MinE now concentrated in that pole
will react until the MinD has diffused back towards the other side and
there is no more membrane bound MinD:ATP to react with.  The cycle
then begins again.

\begin{figure}
  %%\includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-400-50-0-0-1500}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-200-50-0-0-1500}
  \caption{Total protein fluctuation in the right, middle, and left
    parts of a 5$\micron$ by 1$\micron$ pill shaped cell (above) and a
    3$\micron$ by 1$\micron$ pill shaped cell (below).  The vertical axis shows
    stacked the total number of proteins that are of four different
    compound stages and in the right section of the cell (bottom four
    colors), in the middle of the cell (middle four colors) and in the
    left section of the cell (top four colors). The different forms of
    protein changes while the total number of MinD protein in the cell
    remains constant, shown by the fact that the very top line is
    horizontal.}
  \label{box-p}
\end{figure}

Figure \ref{box-p} is a plot of the total fraction of
MinD found in each stage of the cycle in each of three sections of the
cell (we have split the pill into two polar sections and one center
section) over time.  One complete oscillation period is shown, in
which one of the poles begins and ends at a maximum density
concentration and the other a minimum density concentration.  Looking
at the evolution of the latter, we see an initial high concentration
of cytoplasmic MinD:ATP, as the this protein diffuses into the polar
region, followed by a dip in this and large spike in membrane bound
MinD:ATP, as the protein clusters onto the polar cell wall, followed
by a spike in the memebrane bound MinE:MinD:ATP stage of protein,
together with a slight rise in cytoplasmic MinD:ADP as the MinD
dettaches from the wall.

@@

\section{Interpretation of Data}
\section{Conclusion}
\section*{Appendix}
\bibliography{paper}

>>>>

Our diffusion and reaction rates are shown below.  We are interested
primarily in the effect of cellular shape on the protein
oscillations, so we follow Huang's work\cite{huang2003dynamic} and do
not deviate from the wild type values used in that work.  Huang's
simulations use total MinD and MinE concentrations of $1,000/\mu m$
and $350/\mu m$, respectively, in a cylindrical cell of radius $0.5\mu
m$, and in our (non-cylindrical) cells we use the same number of
proteins per unit volume.  These concentration values are 1273 $\mu
m^{-3}$ and 445 $\mu m^{-3}$, respectively.
\begin{gather*}
  \mathcal{D}_D = \mathcal{D}_{E} = 2.5\micron^2/\text{sec}\\
  k_D^{\textrm{ADP $\rightarrow$ ATP}} = 1/\textrm{sec,  }
  k_D = 0.025 \micron /\textrm{sec}\\
  k_{dD} = 0.0015 \micron^3/ \textrm{sec,  }
  k_{de} = 0.7/\textrm{sec}\\
  k_E = 0.093 \micron^3 /\textrm{sec}.
\end{gather*}
A 3D grid is then constructed in cartesian coordinates, with a grid
spacing of .05\micron. We define our cell shapes and solve the
reaction-diffusion partial differential equations numerically to
observe the time evolution of the MinD and MinE concentrations inside
the cell.

@@

We begin with the naturally occuring pill cell shape.  We peice this
shape together as two hemispherical endcaps attached on either end of
a cylinder.  This shape follows the early simulations of Huang
\emph{et al.} but differs in that we have added the end caps for a
more natural shape, expecting similar results.  We test cells of
radius $.5\micron$ and of lengths $2\micron$, $3\micron$, and
$5\micron$ measured from the tip of one end cap to the top of the
other. In each case, we observe quick establishement (within a few
periods) of extremely regular oscillatory behavoir that lasts
indefinitely.  The period increases with cell length, as shown in
Table~\ref{tab:pill-periods}, in agreement with the results of Huang
\emph{et al.}.

\begin{table}
  \begin{tabular}{|r|c|c|c|l|}
    \hline
    Length($\mu$) & 2.50 & 3.00 & 4.00 & 5.00\\
    \hline
    Period(sec) & sim & 33 & 38 & 48 \\ \cline{2-2}
    \hline
  \end{tabular}
  \caption{Period of oscillation according to length of cell for
    pill-shaped cells.}\label{tab:pill-periods}
\end{table}

This simple model is a good starting point for observing in detail the
dynamic interaction between the different stages of protein that lead
to their qualitatively distinct behavoir. Figure \ref{image-p} shows
the process in a series of frames taken from simulation animations.
Each frame is 2.5 seconds ahead of the last, and each image shows a
the concentration of a given state of protein summed over the
coordinate orthogonal to the page.  The color scale for each protein
state is set according to the maximum density observed.

In the beginning, the cytoplasmic MinD:ATP is roughly evenly
distributed throughout the cell with close to its peak density.  There
is a high concentration of MinD on the membrane (in both its
membrane-bound states) at the lower end of the cell, and there is also
cytoplasmic MinD:ADP concentrated in the lower end, which has recently
left the cell membrane.   There is relatively little cytoplasimic
MinE, as it is primarily bound to the lower end of the membrane.

After ??? seconds, very little MinD remains bound to the membrane, and
there is a high concentration of cytoplasmic MinE in the lower half of
the cell.  In the mean time MinD has begun binding in significant
quantities to the upper half of the cell membrane, resulting in a
diminishing of the concentration of cytoplasmic MinD:ATP in the upper
half of the cell.  At this stage, we begin to see the formation of the
``MinE ring'' in the center of the cell---visible here in the plot of
the concentration of membrane-bound MinE:MinD:ATP as two lines
representing a cross section of the ring.  This ring forms as the
cytoplasmic MinE reacts with the membrane-bound MinD:ATP, and leads to
the removal of MinD from the membrane.

From ??? seconds to ??? seconds, we see the MinE ring progress
upwards, as the cytoplasmic MinE diffuses in that direction,
progressively removing the membrane-bound MinD from eventually all but
the very top of the cell membrane.  During this process, the MinD:ATP
bound to the top of the cell membrane is attracting increasing
quantities of MinD:ATP from the cytoplasm.  At ??? seconds, the cell
is in the mirror image state from which it started, and the period is
half-complete.  The second half of the period is a mirror image of the
process just observed, as MinD is driven from the membrane at the top
of the cell, begins binding to the lower half of the cell and a MinE
ring pushes it back down to lower end of the cell.

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-400-50-0-0-1500}
  %\includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-200-50-0-0-1500}
  \caption{Total protein fluctuation in the right, middle, and left
    parts of a 5$\micron$ by 1$\micron$ pill shaped cell (above) and a
    3$\micron$ by 1$\micron$ pill shaped cell (below).  The vertical axis shows
    stacked the total number of proteins that are of four different
    compound stages and in the right section of the cell (bottom four
    colors), in the middle of the cell (middle four colors) and in the
    left section of the cell (top four colors). The different forms of
    protein changes while the total number of MinD protein in the cell
    remains constant, shown by the fact that the very top line is
    horizontal.}
  \label{box-p}
\end{figure}

\davidsays{We should make Fig. \ref{box-p} and Fig. \ref{image-p} plot
  for precisely the same times, that will make it way easier to read
  the two plots.}  While Fig.~\ref{image-p} is helpful for visualizing
the process occuring in the Min system, it is does not show the
relative number of proteins in each state.  To visualize this
information, we plot in Fig.~\ref{box-p} the fraction of MinD found in
each stage of the cycle in each of three sections of the cell---top,
middle and bottom---as a function of time over one period.  The time
axis is the same as in Fig.~\ref{image-p}.  Looking at the evolution
of the latter, we see an initial high concentration of cytoplasmic
MinD:ATP, as the this protein diffuses into the polar region, followed
by a dip in this and large spike in membrane bound MinD:ATP, as the
protein clusters onto the polar cell wall, followed by a spike in the
memebrane bound MinE:MinD:ATP stage of protein, together with a slight
rise in cytoplasmic MinD:ADP as the MinD dettaches from the wall.

@@

\section{Interpretation of Data}
\section{Conclusion}
\bibliography{paper}

\section*{Appendix}
